Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2019-01-10 paper/paper.tex
Use hyper-xr to figure crossref

\documentclass[journal=langd5,email=true, hyperref=true, keywords=false]{achemso}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{todonotes}
\DeclareCaptionLabelSeparator{bar}{  \textbar\quad}
\usepackage[labelsep=bar, labelfont=bf]{caption}
\newcommand{\Fig}{Fig.}
\usepackage{lineno}

@@

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of double layer porous graphene
(PG). Two reservoirs, namely the high-concentration reservoir (HCR)
containing electrolyte solution with molar concentration $c_0$, and
the low-concentration reservoir (LCR) containing de-ionized (DI)
water, are separated by a PG membrane supported by polycarbonate
track-etched (PCTE) film. Magnetic stirrers are used to minimize the
effects of external concentration polarization, and a conductivity
probe is placed in the LCR to monitor the ionic concentration as a
function of time. The membrane is attached to a piece of copper tape
connected to a voltage source applying an electrical bias
$V_{\mathrm{G}}$, with the LCR grounded (\Fig{} \ref{fig:1}a
inset). We fabricated the PCTE-supported PG using the protocol
developed by our group \cite{Choi_2018}, as schematically shown in
\Fig{} \ref{fig:1}b. The normally distributed, circular pores with
pore diameters of 20$\pm$10 nm and a pore density of
1.25$\pm$0.25$\times$10$^{10}$ cm$^{-2}$ (\Fig{} \ref{fig:1}c inset) over a
large area were perforated on graphene.  By optimizing the subsequent
transfer process, a high surface coverage ($>$ 98\%) of PG on PCTE was
obtained (\Fig{} \ref{fig:1}c), allowing reliable characterization of
the ion transport behavior (details see Supplementary 
Section S1).  The system presented here allows us to systematically
investigate the diffusive flux of ions across a sheet of PG,
$\boldsymbol{J}$, driven by the gradient of electrochemical potential
in the electrolyte medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity increases linearly with time in the LCR
(e.g. \Fig{} \ref{fig:2}a) within the measurement duration (1 h),
suggesting the concentration gradient across the membrane remains 
constant. The diffusive fluxes (blue bars in \Fig{} \ref{fig:2}b) of different salts were
obtained by converting the conductivity-time profiles using
calibration curves (Supplementary Section S2). The theoretical diffusive fluxes through PCTE are obtained based on the assumption that the
membrane consists of cylindrical pores with uniform diameter and a low
tortuosity $\tau$ of $\sim{}$1.2 \cite{O_Hern_2012}:
\begin{equation}
  \label{eq:j-pcte}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi r_{\mathrm{PCTE}}^{2} N_{\mathrm{p}} D_{\mathrm{\pm}} \Delta c}{L \tau}
\end{equation}

where $r_{\mathrm{PCTE}}$ is the PCTE pore radius, $N_{\mathrm{p}}$ is
numbers of pore per area, $D_{\mathrm{\pm}}$ is the salt diffusivity
(average of cation and anion diffusivity), $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$ and $L$ is the thickness of the PCTE membrane.
By using the salt diffusivity values (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the transport resistance
of PCTE $R_{\mathrm{PCTE}}$, is dominant over that of PG without gating,
$R_{\mathrm{G}}^{0}$. Using a simple circuit model, we estimate the
ratio $\delta$ between $R_{\mathrm{G}}^{0}$ and $R_{\mathrm{PCTE}}$,
such that $\delta = R_{\mathrm{PCTE}}/R_{\mathrm{G}}^{0}$, to be ~2,
which is in good agreement with the value derived from the
experimental data in \Fig{} \ref{fig:2}b (details see Supplementary Section SXX).
\todo[inline]{Make sure that the values used are correct.}

@@

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the “three-interval” method, which has been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a
illustratively represents how the measured conductivity evolves with
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile with an average slope $s_{1}$ that
corresponds to the diffusive flux across the PG-PCTE membrane. In the
second interval, a voltage $V_{\mathrm{G}}$ is applied to the membrane
after turning off the conductivity probe, followed by the third
interval $s_{3}$, in which the conductivity probe is turned on again
to give the corresponding slope. A control experiment was performed to
ensure the copper tape was not in direct contact with the ionic
solution (Supplementary Section 3). The slope corresponding to the
second interval $s_{2}$, as the conductivity probe is off, is
determined by linear interpolation of the conductivities from the
endpoint of interval 1 to the starting point of interval 3. A decrease
of slope observed in interval 2 indicates that the total flux through
PCTE+PG is reduced upon gating (i.e. salt penetration through the PG
pores is partially rejected). We define such experimentally measured
reduction rate as $\eta$, which is equivalent to the decrease of total
diffusive flux through PCTE+PG as follows:
\begin{equation}
  \label{eq:rejection}
  \eta = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. PG and PCTE membranes effectively form a transport
system with series resistance $R_{\mathrm{PG}}$ (tunable by
$V_{\mathrm{G}}$) and $R_{\mathrm{PCTE}}$ (independent of
$V_{\mathrm{G}}$), respectively. Due to the dominance of
$R_{\mathrm{PCTE}}$ in the ionic transport, significant reduction of
total ionic flux can only be achieved, when
$R_{\mathrm{\mathrm{G}}}(V_{\mathrm{G}} \neq 0) \gg
R_{\mathrm{\mathrm{G}}}^{0}$. Using the circuit analog model, we can
extract the effective salt rejection ratio of PG $\xi$ from the
observable flux reduction ratio $\eta$, as a function of
$V_{\mathrm{G}}$ as (details see Supplementary Section S4):
\begin{equation}
\label{eq:xi-def}
\xi = \frac{(\delta+1) \eta}{\delta \eta + 1}
\end{equation}
The factor $\xi$ enables a reliable and stable measurement of salt
rejection through the PG membrane, as the salt species and initial
conditions may induce a degree of measurement uncertainty between
different samples.

Using 0.1 mM KCl solution in HCR, we measured $\xi$ as a function of
$V_{\mathrm{G}}$ within $\pm1.25$ V, before triggering the
electrochemical reactions (\Fig{} \ref{fig:2}c). We observe an
asymmetric dependence of $\xi$ with respect to $V_{\mathrm{G}}$, with
a higher degree of salt rejection for $V_{\mathrm{G}}>0$. where
positive carriers (holes) are induced in graphene. We notice that an
increase of $\xi$, or a decrease of $\boldsymbol{J}_{\mathrm{PG}}$
with $V_{\mathrm{G}}$, showing an inverse trend compared with those
observed in ionic field effect transistors (IFETs) in which the
diffusive flux increases with the gate voltage
\cite{Nam_2009,Cheng_2018}. A further increase of the KCl
concentration $c_{0}$ influences the attainable degree of salt
rejection. $\xi$ values measured at $V_{\mathrm{G}}$ = 1.25 V,
$\xi_{\mathrm{max}}$, as a function of $c_{0}$ from 10$^{-4}$ to
10$^{-2}$ M KCl (see Supplementary Section S4) exhibit a power law
dependency (\Fig{} \ref{fig:2}d). The $\xi_{\mathrm{max}}-c_{0}$
relation can be nicely fitted by a power law function following
$\xi_{\mathrm{max}} c_{0}^{0.59} = \mathrm{constant}$. As the Debye
length in solution, $\lambda_{\mathrm{D}}$, is inversely proportional
to $c^{0.5}$, we infer that the $V_{\mathrm{G}}$-induced salt
rejection originates from the modulation of the electrical double
layer (EDL), as will be discussed later.

@@

$\Delta \phi_{\mathrm{G}} = \phi_{\mathrm{G}}(V_{\mathrm{g}}) -
\phi_{\mathrm{G}}(V_{\mathrm{g}}=0)$ is the change of graphene’s work
function. The properties of the PG samples fabricated is close to
intrinsic graphene, with charge density at $V_{\mathrm{G}}=0$ measured
to be 1.4$\times$10$^{11}$ \textit{e}$\cdot$cm$^{-2}$, corresponding
to a work function difference of 48 meV from graphene's Dirac point
(details see Supplementary Section S5), much smaller than the
$\Delta \phi_{\mathrm{G}}$ induced by gating.  Therefore we adapt the
simplification that the charge neutrality point of graphene coincides
with the Dirac point at $V_{\mathrm{G}}$ = 0. The elementary
electronic properties of graphene give:

@@

where $\hbar$ is the reduced Planck constant, and
$v_{\mathrm{F}}=1.1\times10^{6}$ m$\cdot$s$^{-1}$ is the Fermi
velocity of graphene (more details find Supplementary Section S6).
Note that the PG film fabricated experimentally is double-layer
turbostratic (randomly aligned) graphene, in which we assume the
$\Delta \phi_{\mathrm{G}} - \sigma_{\mathrm{G}}$ dependence follows
Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing
Eqs. \eqref{eq:pnp}-\eqref{eq:poisson} using the finite-element method
(FEM), in which the graphene surface potential is coupled with
Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved
simultaneously to reach convergent numerical solutions of
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ for a given
$V_{\mathrm{G}}$. Clearly, the above model yields symmetric
characteristics for $\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ with
respect to $V_{\mathrm{G}}$ due to the symmetric band structure of
graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene
synthesized by chemical vapor deposition (CVD) used in our
experiments, it is well-recognized that the electron traps are
inherently introduced during synthesis and patterning
\cite{Dean_2010}, effectively reducing the charge density and surface
potential on graphene $V_{\mathrm{G}}<0$ (Supplementary \Fig{} S3),
which is related to the fact that observed $\xi$ is much lower within
the $V_{\mathrm{G}}<0$ regime in \Fig{} \ref{fig:2}c. With the
nonideality in mind, hereafter, we compare the calculations and
experiments in the regime of $V_{\mathrm{G}}>0$. For example, \Fig{}
\ref{fig:3}c presents the calculated $\sigma_{\mathrm{G}}$ and
$\psi_{\mathrm{G}}$ as a function of $V_{\mathrm{G}}$ considering the
ionic diffusion through a single 20 nm-diameter nanopore drilled on a
sheet of semi-infinite double layer graphene that separates HCR
containing KCl solution at $c_{0}$ = 0.1 mM, and LCR at $c_{0}/10$ in
axisymmetric coordinates (details see Supplementary Section
S5). Indeed, the interplay between graphene quantum capacitance and
the EDL significantly reduces the graphene surface potential
$\psi_{\mathrm{G}}$ to $\sim$0.3 V at $V_{\mathrm{G}}$ = 1.25 V,
corresponding a surface charge density $\sigma_{\mathrm{G}}$ of
$\sim$0.08 C$\cdot$m$^{-2}$. Note that as we mentioned earlier, an
important merit of the system here is that the surface charge density
can be precisely determined and controlled, rather than being treated
as a fitting parameter as in most of the 2D nanopore studies
\cite{Rollings_2016}.

@@

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes
small. (also see Supplementary \Fig{} S4). We also observe that
$\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
S5). The observations confirm the distinct ionic transport pathways
upon applying a positive $V_{\mathrm{G}}$: pore center for cations and
pore edge for anions. \Fig{} \ref{fig:4}c presents the calculated
\textit{z}-component of the cation and anion fluxes through the pore
at the graphene plane (z = 0), $\boldsymbol{J}_{z}$ ,as a function of
the normalized radius, $r/r_{\mathrm{G}}$ , where $r$ is the radial
coordinate and $r_{\mathrm{G}}$ is the pore radius, at different
$V_{\mathrm{G}}$ values. Note that $\boldsymbol{J}_{z}$ is negative
because the LCR is placed at z$<$0 in the simulation box. At
$V_{\mathrm{G}}$ = 0 (pure diffusion), the anion and cation flux
profiles are identical, with a higher flux near the pore edge, which
is expected, as the concentration gradient is higher. By gradually
increasing $V_{\mathrm{G}}$, both fluxes are reduced throughout the
pore, while the degree of reduction is more pronounced at the pore
edge for the cation, and at the pore center for the anion, following
the scenarios we discussed above. Accordingly, the integrated fluxes
across the pore, $|\boldsymbol{J}_{z}|$, are reduced with
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:4}d). A total flux reduction of up
to 80\% is predicted. Another interesting finding here is that, by
increasing $V_{\mathrm{G}}$, the nanopore transport preferentially
allows cations over anions, or in other words, the anion flux is more
reduced with $V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the
ion selectivity of graphene nanopore \cite{Rollings_2016}. In our setup, the imbalance between cation and anion
fluxes near the graphene nanopore is compensated by the EDL formation
at the counter-electrode in LCR, which maintains the electroneutrality
of the system. Notably, since atomically thin graphene samples are
used, the mechanism of salt rejection is clearly distinct from that in
capacitive deionization \cite{Biesheuvel_2010}, where electrodes with large
surface area are required.

>>>>

\documentclass[journal=langd5,email=true, hyperref=true, keywords=false]{achemso}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{todonotes}
\DeclareCaptionLabelSeparator{bar}{  \textbar\quad}
\usepackage[labelsep=bar, labelfont=bf]{caption}
\newcommand{\Fig}{Fig.}
\usepackage{lineno}
\usepackage{xr-hyper}
\externaldocument[SI-]{SI}

@@

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of double layer porous graphene
(PG). Two reservoirs, namely the high-concentration reservoir (HCR)
containing electrolyte solution with molar concentration $c_0$, and
the low-concentration reservoir (LCR) containing de-ionized (DI)
water, are separated by a PG membrane supported by polycarbonate
track-etched (PCTE) film. Magnetic stirrers are used to minimize the
effects of external concentration polarization, and a conductivity
probe is placed in the LCR to monitor the ionic concentration as a
function of time. The membrane is attached to a piece of copper tape
connected to a voltage source applying an electrical bias
$V_{\mathrm{G}}$, with the LCR grounded (\Fig{} \ref{fig:1}a
inset). We fabricated the PCTE-supported PG using the protocol
developed by our group \cite{Choi_2018}, as schematically shown in
\Fig{} \ref{fig:1}b. The normally distributed, circular pores with
pore diameters of 20$\pm$10 nm and a pore density of
1.25$\pm$0.25$\times$10$^{10}$ cm$^{-2}$ (\Fig{} \ref{fig:1}c inset) over a
large area were perforated on graphene.  By optimizing the subsequent
transfer process, a high surface coverage ($>$ 98\%) of PG on PCTE was
obtained (\Fig{} \ref{fig:1}c), allowing reliable characterization of
the ion transport behavior (details see Supplementary 
Section \ref{SI-sec:exp}).  The system presented here allows us to systematically
investigate the diffusive flux of ions across a sheet of PG,
$\boldsymbol{J}$, driven by the gradient of electrochemical potential
in the electrolyte medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity increases linearly with
time in the LCR (e.g. \Fig{} \ref{fig:2}a) within the measurement
duration (1 h), suggesting the concentration gradient across the
membrane remains constant. The diffusive fluxes (blue bars in \Fig{}
\ref{fig:2}b) of different salts were obtained by converting the
conductivity-time profiles using calibration curves (Supplementary
Section \ref{SI-sec:measure}). The theoretical diffusive fluxes
through PCTE are obtained based on the assumption that the membrane
consists of cylindrical pores with uniform diameter and a low
tortuosity $\tau$ of $\sim{}$1.2 \cite{O_Hern_2012}:
\begin{equation}
  \label{eq:j-pcte}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi r_{\mathrm{PCTE}}^{2} N_{\mathrm{p}} D_{\mathrm{\pm}} \Delta c}{L \tau}
\end{equation}

where $r_{\mathrm{PCTE}}$ is the PCTE pore radius, $N_{\mathrm{p}}$ is
numbers of pore per area, $D_{\mathrm{\pm}}$ is the salt diffusivity
(average of cation and anion diffusivity), $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$ and $L$ is the thickness of the PCTE membrane.
By using the salt diffusivity values (Supplementary Table
\ref{SI-tab:diff}), the calculated diffusive fluxes (red bars in
\Fig{} \ref{fig:2}b) show reasonable agreement with the experimental
values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the transport
resistance of PCTE $R_{\mathrm{PCTE}}$, is dominant over that of PG
without gating, $R_{\mathrm{G}}^{0}$. Using a simple circuit model, we
estimate the ratio $\delta$ between $R_{\mathrm{G}}^{0}$ and
$R_{\mathrm{PCTE}}$, such that
$\delta = R_{\mathrm{PCTE}}/R_{\mathrm{G}}^{0}$, to be ~2, which is in
good agreement with the value derived from the experimental data in
\Fig{} \ref{fig:2}b (details see Supplementary Section
\ref{SI-sec:R-model}).
\todo[inline]{Make sure that the values used
  are correct.}

@@

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the “three-interval” method, which has been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a
illustratively represents how the measured conductivity evolves with
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile with an average slope $s_{1}$ that
corresponds to the diffusive flux across the PG-PCTE membrane. In the
second interval, a voltage $V_{\mathrm{G}}$ is applied to the membrane
after turning off the conductivity probe, followed by the third
interval $s_{3}$, in which the conductivity probe is turned on again
to give the corresponding slope. A control experiment was performed to
ensure the copper tape was not in direct contact with the ionic
solution (Supplementary Section \ref{SI-sec:copper}). The slope corresponding to the
second interval $s_{2}$, as the conductivity probe is off, is
determined by linear interpolation of the conductivities from the
endpoint of interval 1 to the starting point of interval 3. A decrease
of slope observed in interval 2 indicates that the total flux through
PCTE+PG is reduced upon gating (i.e. salt penetration through the PG
pores is partially rejected). We define such experimentally measured
reduction rate as $\eta$, which is equivalent to the decrease of total
diffusive flux through PCTE+PG as follows:
\begin{equation}
  \label{eq:rejection}
  \eta = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. PG and PCTE membranes effectively form a transport
system with series resistance $R_{\mathrm{PG}}$ (tunable by
$V_{\mathrm{G}}$) and $R_{\mathrm{PCTE}}$ (independent of
$V_{\mathrm{G}}$), respectively. Due to the dominance of
$R_{\mathrm{PCTE}}$ in the ionic transport, significant reduction of
total ionic flux can only be achieved, when
$R_{\mathrm{\mathrm{G}}}(V_{\mathrm{G}} \neq 0) \gg
R_{\mathrm{\mathrm{G}}}^{0}$. Using the circuit analog model, we can
extract the effective salt rejection ratio of PG $\xi$ from the
observable flux reduction ratio $\eta$, as a function of
$V_{\mathrm{G}}$ as (details see Supplementary Section
\ref{SI-sec:R-model}):
\begin{equation}
\label{eq:xi-def}
\xi = \frac{(\delta+1) \eta}{\delta \eta + 1}
\end{equation}
The factor $\xi$ enables a reliable and stable measurement of salt
rejection through the PG membrane, as the salt species and initial
conditions may induce a degree of measurement uncertainty between
different samples.

Using 0.1 mM KCl solution in HCR, we measured $\xi$ as a function of
$V_{\mathrm{G}}$ within $\pm1.25$ V, before triggering the
electrochemical reactions (\Fig{} \ref{fig:2}c). We observe an
asymmetric dependence of $\xi$ with respect to $V_{\mathrm{G}}$, with
a higher degree of salt rejection for $V_{\mathrm{G}}>0$. where
positive carriers (holes) are induced in graphene. We notice that an
increase of $\xi$, or a decrease of $\boldsymbol{J}_{\mathrm{PG}}$
with $V_{\mathrm{G}}$, showing an inverse trend compared with those
observed in ionic field effect transistors (IFETs) in which the
diffusive flux increases with the gate voltage
\cite{Nam_2009,Cheng_2018}. A further increase of the KCl
concentration $c_{0}$ influences the attainable degree of salt
rejection. $\xi$ values measured at $V_{\mathrm{G}}$ = 1.25 V,
$\xi_{\mathrm{max}}$, as a function of $c_{0}$ from 10$^{-4}$ to
10$^{-2}$ M KCl (see Supplementary Section \ref{SI-sec:debye}) exhibit a power law
dependency (\Fig{} \ref{fig:2}d). The $\xi_{\mathrm{max}}-c_{0}$
relation can be nicely fitted by a power law function following
$\xi_{\mathrm{max}} c_{0}^{0.59} = \mathrm{constant}$. As the Debye
length in solution, $\lambda_{\mathrm{D}}$, is inversely proportional
to $c^{0.5}$, we infer that the $V_{\mathrm{G}}$-induced salt
rejection originates from the modulation of the electrical double
layer (EDL), as will be discussed later.

@@

$\Delta \phi_{\mathrm{G}} = \phi_{\mathrm{G}}(V_{\mathrm{g}}) -
\phi_{\mathrm{G}}(V_{\mathrm{g}}=0)$ is the change of graphene’s work
function. The properties of the PG samples fabricated is close to
intrinsic graphene, with charge density at $V_{\mathrm{G}}=0$ measured
to be 1.4$\times$10$^{11}$ \textit{e}$\cdot$cm$^{-2}$, corresponding
to a work function difference of 48 meV from graphene's Dirac point
(details see Supplementary Section \ref{SI-sec:charge-dens}), much
smaller than the $\Delta \phi_{\mathrm{G}}$ induced by gating.
Therefore we adapt the simplification that the charge neutrality point
of graphene coincides with the Dirac point at $V_{\mathrm{G}}$ =
0. The elementary electronic properties of graphene give:

@@

where $\hbar$ is the reduced Planck constant, and
$v_{\mathrm{F}}=1.1\times10^{6}$ m$\cdot$s$^{-1}$ is the Fermi
velocity of graphene (more details find Supplementary Section
\ref{SI-sec:charge-dens}).  Note that the PG film fabricated
experimentally is double-layer turbostratic (randomly aligned)
graphene, in which we assume the
$\Delta \phi_{\mathrm{G}} - \sigma_{\mathrm{G}}$ dependence follows
Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing
Eqs. \eqref{eq:pnp}-\eqref{eq:poisson} using the finite-element method
(FEM), in which the graphene surface potential is coupled with
Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved
simultaneously to reach convergent numerical solutions of
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ for a given
$V_{\mathrm{G}}$. Clearly, the above model yields symmetric
characteristics for $\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ with
respect to $V_{\mathrm{G}}$ due to the symmetric band structure of
graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene
synthesized by chemical vapor deposition (CVD) used in our
experiments, it is well-recognized that the electron traps are
inherently introduced during synthesis and patterning
\cite{Dean_2010}, effectively reducing the charge density and surface
potential on graphene $V_{\mathrm{G}}<0$ (Supplementary \Fig{}
\ref{SI-fig:trap}), which is related to the fact that observed $\xi$
is much lower within the $V_{\mathrm{G}}<0$ regime in \Fig{}
\ref{fig:2}c. With the nonideality in mind, hereafter, we compare the
calculations and experiments in the regime of $V_{\mathrm{G}}>0$. For
example, \Fig{} \ref{fig:3}c presents the calculated
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ as a function of
$V_{\mathrm{G}}$ considering the ionic diffusion through a single 20
nm-diameter nanopore drilled on a sheet of semi-infinite double layer
graphene that separates HCR containing KCl solution at $c_{0}$ = 0.1
mM, and LCR at $c_{0}/10$ in axisymmetric coordinates (details see
Supplementary Section \ref{SI-fig:scheme}). Indeed, the interplay between graphene
quantum capacitance and the EDL significantly reduces the graphene
surface potential $\psi_{\mathrm{G}}$ to $\sim$0.3 V at
$V_{\mathrm{G}}$ = 1.25 V, corresponding a surface charge density
$\sigma_{\mathrm{G}}$ of $\sim$0.08 C$\cdot$m$^{-2}$. Note that as we
mentioned earlier, an important merit of the system here is that the
surface charge density can be precisely determined and controlled,
rather than being treated as a fitting parameter as in most of the 2D
nanopore studies \cite{Rollings_2016}.

@@

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes small (see
Supplementary \Fig{} \ref{SI-fig:conc}). We also observe that
$\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
\ref{SI-fig:potential}). The observations confirm the distinct ionic
transport pathways upon applying a positive $V_{\mathrm{G}}$: pore
center for cations and pore edge for anions. \Fig{} \ref{fig:4}c
presents the calculated \textit{z}-component of the cation and anion
fluxes through the pore at the graphene plane (z = 0),
$\boldsymbol{J}_{z}$ ,as a function of the normalized radius,
$r/r_{\mathrm{G}}$ , where $r$ is the radial coordinate and
$r_{\mathrm{G}}$ is the pore radius, at different $V_{\mathrm{G}}$
values. Note that $\boldsymbol{J}_{z}$ is negative because the LCR is
placed at z$<$0 in the simulation box. At $V_{\mathrm{G}}$ = 0 (pure
diffusion), the anion and cation flux profiles are identical, with a
higher flux near the pore edge, which is expected, as the
concentration gradient is higher. By gradually increasing
$V_{\mathrm{G}}$, both fluxes are reduced throughout the pore, while
the degree of reduction is more pronounced at the pore edge for the
cation, and at the pore center for the anion, following the scenarios
we discussed above. Accordingly, the integrated fluxes across the
pore, $|\boldsymbol{J}_{z}|$, are reduced with $V_{\mathrm{G}}$
(\Fig{} \ref{fig:4}d). A total flux reduction of up to 80\% is
predicted. Another interesting finding here is that, by increasing
$V_{\mathrm{G}}$, the nanopore transport preferentially allows cations
over anions, or in other words, the anion flux is more reduced with
$V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the ion
selectivity of graphene nanopore \cite{Rollings_2016}. In our setup,
the imbalance between cation and anion fluxes near the graphene
nanopore is compensated by the EDL formation at the counter-electrode
in LCR, which maintains the electroneutrality of the system. Notably,
since atomically thin graphene samples are used, the mechanism of salt
rejection is clearly distinct from that in capacitive deionization
\cite{Biesheuvel_2010}, where electrodes with large surface area are
required.

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the ionic
transport through an IFET, in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity \cite{Nam_2009,Lee_2015,Feng_2016}. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (details see Supplementary Section
\ref{SI-sec:simu-res}). Intriguingly, by increasing $\psi_{\mathrm{G}}$
larger than 400 mV, the ionic flux starts to increase, reversing the
trend observed at the low $\psi_{\mathrm{G}}$ regime, due to (i) the
drift flux become dominant over diffusion flux and (ii) the
concentration near the pore edge (prefered path) greatly increases
(details see Supplementary Section \ref{SI-sec:simu-res}, Supplementary
Figs. \ref{SI-fig:reverse} and \ref{SI-fig:large-V}). This level of
$\psi_{\mathrm{G}}$ cannot be reached experimentally in graphene
membrane, as this requires a $V_{\mathrm{G}}$ larger than 2.0 V, which
triggers the electrochemical reactions \cite{Toh_2011}. On the other
hand, in an IFET, the electric potential on the pore wall is
considerably higher, equivalent to $V_{\mathrm{G}}$ due to the
metallic nature of gate electrode (having an infinitely-large quantum
capacitance). We therefore conclude that the quantum
capacitance-induced nonlinear damping in the surface potential results
in the observed salt rejection.  Following the above discussions, we
further investigate the physical limits for the biased graphene
nanopores. Specifically, the salt rejection characteristics are
controlled by: (i) the overlap of EDL inside the nanopore, essentially
controlled by two length scales of $\lambda_{\mathrm{D}}$ and
$r_{\mathrm{G}}$, and (ii) the graphene surface potential
$\psi_{\mathrm{G}}$ controlled by $V_{\mathrm{G}}$ following
Eq. \eqref{eq:Vg}. To this end, we calculate $\psi_{\mathrm{G}}$ as a
function of $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:5}a). Clearly, when the bulk
concentration $c_{0}$ increases, a higher $V_{\mathrm{G}}$ is required
to reach the same level of $\psi_{\mathrm{G}}$. Consequently, as
revealed in \Fig{} \ref{fig:5}b, the calculated salt rejection factor
$\xi$ increases with both $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$, in line with the experimentally observed
$\xi - c_{0}$ dependence. We also find the overall trends for
$\psi_{\mathrm{G}}$ and $\xi$ with respect to
$\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$
= 1.0, a high value of $\xi$ of $>$ 0.85 can be achieved.
